---
title: "House_Prices"
author: "Don Padmaperuma"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library('ggplot2')
library('ggthemes') 
library('scales')
library('dplyr') 
library('mice')
library('randomForest') 
library('data.table')
library('gridExtra')
library('corrplot') 
library('GGally')
library('e1071')
library(caret)
```

## Acquire Data

```{r}
train <-read.csv('train.csv', stringsAsFactors = F)
test  <-read.csv('test.csv', stringsAsFactors = F)
```

## Explore Data

```{r}
dim(train)
```
```{r}
str(train)
```
```{r}
dim(test)
```

```{r}
str(test)
```

```{r}
#Count the number of columns that consists of text data

sum(sapply(train[,1:81], typeof) == "character")
```

```{r}
#Count the number of columns that consists of numerical data

sum(sapply(train[,1:81], typeof) == "integer")
```
```{r}
# Obtain summary statistics

summary(train[,sapply(train[,1:81], typeof) == "integer"])
```
```{r}
# The percentage of data missing in train

sum(is.na(train)) / (nrow(train) *ncol(train))
```
```{r}
# The percentage of data missing in test

sum(is.na(test)) / (nrow(test) * ncol(test))
```

```{r}
# Check for duplicated rows

cat("The number of duplicated rows are", nrow(train) - nrow(unique(train)))
```

## Combine data
Since test dataset has no “Saleprice” variable. We will create it and then combine.

```{r}
test$SalePrice<-rep(NA,1459)
house<-bind_rows(train,test)

## Data Exploration

str(house)
```
## Data Visualization

```{r}
cat_var <- names(train)[which(sapply(train, is.character))]
cat_car <- c(cat_var, 'BedroomAbvGr', 'HalfBath', ' KitchenAbvGr','BsmtFullBath', 'BsmtHalfBath', 'MSSubClass')
numeric_var <- names(train)[which(sapply(train, is.numeric))]
## Creating one training dataset with categorical variable and one with numeric variable. We will use this for data visualization.

train1_cat<-train[cat_var]
train1_num<-train[numeric_var]
```

### Bar plot/Density plot function
```{r}
## Bar plot function 

plotHist <- function(data_in, i) 
{
  data <- data.frame(x=data_in[[i]])
  p <- ggplot(data=data, aes(x=factor(x))) + stat_count() + xlab(colnames(data_in)[i]) + theme_light() + 
    theme(axis.text.x = element_text(angle = 90, hjust =1))
  return (p)
}
## Density plot function

plotDen <- function(data_in, i){
  data <- data.frame(x=data_in[[i]], SalePrice = data_in$SalePrice)
  p <- ggplot(data= data) + geom_line(aes(x = x), stat = 'density', size = 1,alpha = 1.0) +
    xlab(paste0((colnames(data_in)[i]), '\n', 'Skewness: ',round(skewness(data_in[[i]], na.rm = TRUE), 2))) + theme_light() 
  return(p)
}
```

### Function to call both Bar plot and Density plot function
```{r}
doPlots <- function(data_in, fun, ii, ncol=3) 
{
  pp <- list()
  for (i in ii) {
    p <- fun(data_in=data_in, i=i)
    pp <- c(pp, list(p))
  }
  do.call("grid.arrange", c(pp, ncol=ncol))
}


## Barplots for the categorical features

doPlots(train1_cat, fun = plotHist, ii = 1:4, ncol = 2)
```

### Box Plot
```{r}
ggplot(train, aes(x = Neighborhood, y = SalePrice)) +
  geom_boxplot() +
  geom_hline(aes(yintercept=80), 
             colour='red', linetype='dashed', lwd=2) +
  scale_y_continuous(labels=dollar_format()) +
  theme_few()
```
## Missing Value 
```{r}
# Looking at the distribution and summary of the target variable

summary(train$SalePrice)
```
```{r}
quantile(train$SalePrice)
```

```{r}
## Histogram for target variable

hist(train$SalePrice)
```

```{r}
# Plotting 'GrLivArea' too see if there are any outliers

ggplot(train,aes(y=SalePrice,x=GrLivArea))+geom_point()
```

```{r}
summary(train$GrLivArea)
```
```{r}
# There are outliers in 'GrLivArea' field. Let's remove those outliers.

train <- train[train$GrLivArea<=4000,]
```

```{r}
## To find number of missing value for all variable in train dataset

colSums(sapply(train, is.na))
```
```{r}
## To find number of missing value for all variable in combined dataset (Train+Test)

sapply(house[,1:80], function(x) sum(is.na(x)))
```

```{r}
## Taking all the missing data indices in one variables

Missing_indices <- sapply(train,function(x) sum(is.na(x)))
Missing_Summary <- data.frame(index = names(train),Missing_Values=Missing_indices)
Missing_Summary[Missing_Summary$Missing_Values > 0,]
```

```{r}
## Combining train and test data for quicker data prep
test$SalePrice <- NA
train$isTrain <- 1
test$isTrain <- 0
house <- rbind(train,test)
```

```{r}
house$MasVnrArea[which(is.na(house$MasVnrArea))] <- mean(house$MasVnrArea,na.rm=T)
```

### Changing all the NA in variables to None 
```{r}
house$Alley1 <- as.character(house$Alley)
house$Alley1[which(is.na(house$Alley))] <- "None"
table(house$Alley1)
```
```{r}
house$Alley <- as.factor(house$Alley1)
house <- subset(house,select = -Alley1)
```

```{r}
house$MasVnrType1 <- as.character(house$MasVnrType)
house$MasVnrType1[which(is.na(house$MasVnrType))] <- "None"
house$MasVnrType <- as.factor(house$MasVnrType1)
house <- subset(house,select = -MasVnrType1)
table(house$MasVnrType)
```
```{r}
house$LotFrontage[which(is.na(house$LotFrontage))] <- median(house$LotFrontage,na.rm = T)
```

```{r}
house$FireplaceQu1 <- as.character(house$FireplaceQu)
house$FireplaceQu1[which(is.na(house$FireplaceQu))] <- "None"
house$FireplaceQu <- as.factor(house$FireplaceQu1)
house <- subset(house,select = -FireplaceQu1)
```

```{r}
house$PoolQC1 <- as.character(house$PoolQC)
house$PoolQC1[which(is.na(house$PoolQC))] <- "None"
house$PoolQC <- as.factor(house$PoolQC1)
house <- subset(house,select = -PoolQC1)
```

```{r}
house$Fence1 <- as.character(house$Fence)
house$Fence1[which(is.na(house$Fence))] <- "None"
house$Fence <- as.factor(house$Fence1)
house <- subset(house,select = -Fence1)
```

```{r}
house$MiscFeature1 <- as.character(house$MiscFeature)
house$MiscFeature1[which(is.na(house$MiscFeature))] <- "None"
house$MiscFeature <- as.factor(house$MiscFeature1)
house <- subset(house,select = -MiscFeature1)
```

```{r}
house$GarageType1 <- as.character(house$GarageType)
house$GarageType1[which(is.na(house$GarageType))] <- "None"
house$GarageType <- as.factor(house$GarageType1)
house <- subset(house,select = -GarageType1)
```

```{r}
house$GarageYrBlt[which(is.na(house$GarageYrBlt))] <- 0
```

```{r}
house$GarageFinish1 <- as.character(house$GarageFinish)
house$GarageFinish1[which(is.na(house$GarageFinish))] <- "None"
house$GarageFinish <- as.factor(house$GarageFinish1)
house <- subset(house,select = -GarageFinish1)

```

```{r}
house$GarageQual1 <- as.character(house$GarageQual)
house$GarageQual1[which(is.na(house$GarageQual))] <- "None"
house$GarageQual <- as.factor(house$GarageQual1)
house <- subset(house,select = -GarageQual1)
```

```{r}
house$GarageCond1 <- as.character(house$GarageCond)
house$GarageCond1[which(is.na(house$GarageCond))] <- "None"
house$GarageCond <- as.factor(house$GarageCond1)
house <- subset(house,select = -GarageCond1)
```

```{r}
house$BsmtQual1 <- as.character(house$BsmtQual)
house$BsmtQual1[which(is.na(house$BsmtQual))] <- "None"
house$BsmtQual <- as.factor(house$BsmtQual1)
house <- subset(house,select = -BsmtQual1)
```

```{r}
house$BsmtCond1 <- as.character(house$BsmtCond)
house$BsmtCond1[which(is.na(house$BsmtCond))] <- "None"
house$BsmtCond <- as.factor(house$BsmtCond1)
house <- subset(house,select = -BsmtCond1)

```

```{r}

house$BsmtExposure1 <- as.character(house$BsmtExposure)
house$BsmtExposure1[which(is.na(house$BsmtExposure))] <- "None"
house$BsmtExposure <- as.factor(house$BsmtExposure1)
house <- subset(house,select = -BsmtExposure1)
```

```{r}
house$BsmtFinType11 <- as.character(house$BsmtFinType1)
house$BsmtFinType11[which(is.na(house$BsmtFinType1))] <- "None"
house$BsmtFinType1 <- as.factor(house$BsmtFinType11)
house <- subset(house,select = -BsmtFinType11)
```

```{r}
house$BsmtFinType21 <- as.character(house$BsmtFinType2)
house$BsmtFinType21[which(is.na(house$BsmtFinType2))] <- "None"
house$BsmtFinType2 <- as.factor(house$BsmtFinType21)
house <- subset(house,select = -BsmtFinType21)
```

```{r}
house$Electrical1 <- as.character(house$Electrical)
house$Electrical1[which(is.na(house$Electrical))] <- "None"
house$Electrical <- as.factor(house$Electrical1)
house <- subset(house,select = -Electrical1)
```

### Factorizing
```{r}
house$MSZoning<- factor(house$MSZoning)
house$Street <- factor(house$Street)
house$LotShape <-factor(house$LotShape )
house$LandContour<-factor(house$LandContour)
house$Utilities<-factor(house$Utilities)
house$LotConfig<-factor(house$LotConfig)
house$LandSlope<-factor(house$LandSlope)
house$Neighborhood<-factor(house$Neighborhood)
house$Condition1<-factor(house$Condition1)
house$Condition2<-factor(house$Condition2)
house$BldgType<-factor(house$BldgType)
house$HouseStyle<-factor(house$HouseStyle)
house$RoofStyle<-factor(house$RoofStyle)
house$RoofMatl<-factor(house$RoofMatl)
house$Exterior1st<-factor(house$Exterior1st)
house$Exterior2nd<-factor(house$Exterior2nd)
house$ExterQual<-factor(house$ExterQual)
house$ExterCond<-factor(house$ExterCond)
house$Foundation<-factor(house$Foundation)
house$Heating<-factor(house$Heating)
house$HeatingQC<-factor(house$HeatingQC)
house$CentralAir<-factor(house$CentralAir)
house$KitchenQual<-factor(house$KitchenQual)
house$Functional<-factor(house$Functional)
house$PavedDrive<-factor(house$PavedDrive)
house$SaleType<-factor(house$SaleType)
house$SaleCondition<-factor(house$SaleCondition)
str(house)
```
### Taking all the column classes in one variable so as to seperate factors from numerical variables.

```{r}
Column_classes <- sapply(names(house),function(x){class(house[[x]])})
numeric_columns <-names(Column_classes[Column_classes != "factor"])

#determining skew of each numeric variable

skew <- sapply(numeric_columns,function(x){skewness(house[[x]],na.rm = T)})

# Let us determine a threshold skewness and transform all variables above the treshold.

skew <- skew[skew > 0.75]

# transform excessively skewed features with log(x + 1)

for(x in names(skew)) 
{
  house[[x]] <- log(house[[x]] + 1)
}
```

### Train and test dataset creation

```{r}
train <- house[house$isTrain==1,]
test <- house[house$isTrain==0,]
smp_size <- floor(0.75 * nrow(train))

## setting the seed to make the partition reproducible

set.seed(123)
train_ind <- sample(seq_len(nrow(train)), size = smp_size)

train_new <- train[train_ind, ]
validate <- train[-train_ind, ]
train_new <- subset(train_new,select=-c(Id,isTrain))
validate <- subset(validate,select=-c(Id,isTrain))
nrow(train_new)
```


### Build the model 

```{r}
library(randomForest)
house_model <- randomForest(SalePrice~.,
                            data = train_new)
```

### Get importance
```{r}
importance    <- importance(house_model)
varImpPlot(house_model)
```

```{r}
# Predict using the test set

prediction <- predict(house_model,test)

# Evaluation RMSE function

RMSE <- function(x,y){
  a <- sqrt(sum((log(x)-log(y))^2)/length(y))
  return(a)
}
```

```{r}
RMSE1 <- RMSE(prediction, validate$SalePrice)
RMSE1
```

```{r}
prediction[which(is.na(prediction))] <- mean(prediction,na.rm=T)
submit <- data.frame(Id=test$Id,SalePrice=prediction)
write.csv(submit,file="my_submission.csv",row.names=F)
```




























